Online {\IterComp}

### 2017-09-15 motivation.tex
overall changes

    \begin{figure}[t!]
        \centering
        \includegraphics[width=0.5\textwidth]{figs/motivation-online.pdf}
        \caption{
            Histograms of speedup over \texttt{-O3} for each of the 1,000 inputs of \texttt{susan\_c}. The green histogram shows the
            speedup when using the optimization sequence selected by offline \itercomp, the blue one is the speedup using the optimization
            sequence selected by an online-like approach. \FIXME{Beware Daltonism!}
        }
        \label{fig:motivation-online}
    \end{figure}

    We performed offline iterative compilation on \texttt{susan\_c} to find the optimization settings that maximize the average performance
    over five randomly chosen inputs from a set of 1,000 provided by the \textit{KDataSets} benchmark suite~\cite{chen10,chen12a}.
    The green histogram in Figure~\ref{fig:motivation-online} shows for each of the 1,0000 inputs the speedup versus \texttt{-O3} of that
    chosen optimization setting. The dark green line shows the average speedup for all those inputs. The five training inputs are
    all among the top performing inputs. On the other hand, the chosen optimization does poorly on other inputs, often causing
    a slowdown and yielding only a slight speedup of 4\% on average.

@@

    Prior to this paper there were two proposed metrics to correlate with program efficiency. The first is simply to use runtime, ignoring
    the work aspect. This functions well when the inputs all require similar effort. When that is not the case it bears little or no
    correlation to efficiency. The second approach is \textit{instructions per cycle}~(IPC), suggested by~\citep{fursin07}. While IPC seems
    promising in the sense that faster instruction execution is a good thing, this is belied by the simple observation that adding
    redundant NOPs to a program will increase IPC while also slowing the program down. Figure~\ref{fig:motivation-metric} shows the
    relationship between execution time, IPC and speedup for different optimizations of \texttt{susan\_c}, averaged across a subset of the
    inputs. Neither IPC nor execution time correlate well, and so neither can be used for iterative compilation. We need a work efficiency
    metric which can approximate speedup instead, using only a single execution.

>>>>

    \begin{figure}[t!]
        \centering
        \includegraphics[width=0.5\textwidth]{figs/motivation-online.pdf}
        \caption{
            Histograms of speedup over \texttt{-O3} for each of the 1,000 inputs of \texttt{susan\_c}. The green histogram shows the
            speedup when using the optimization sequence selected by offline \itercomp, the blue one is the speedup using the optimization
            sequence selected by an online-like approach.
        }
        \label{fig:motivation-online}
    \end{figure}

    We performed offline iterative compilation on the \texttt{susan\_c} benchmark to find the optimization settings that maximize the average performance
    over five randomly chosen inputs from a set of 1,000 provided by the \textit{KDataSets} benchmark suite~\cite{chen10,chen12a}.
    The green histogram in Figure~\ref{fig:motivation-online} shows, for each of the 1,000 inputs, the speedup versus \texttt{-O3} of that
    chosen optimization setting. The dark green line shows the average speedup for all those inputs. The five training inputs are
    all among the top performing inputs. On the other hand, the chosen optimization does poorly on other inputs, often causing
    a slowdown and yielding only a slight speedup of 4\% on average.

@@

    Prior to this paper there were two proposed metrics to correlate with program efficiency. The first is simply to use runtime, ignoring
    the work aspect. This functions well when the inputs all require similar effort. When that is not the case it bears little or no
    correlation to efficiency. The second approach is \textit{instructions per cycle}~(IPC), suggested by~\citep{fursin07}. While IPC seems
    promising in the sense that faster instruction execution is a good thing, this is belied by the simple observation that adding
    redundant NOPs to a program will increase IPC while also slowing the program down. Figure~\ref{fig:motivation-metric} shows the
    relationship between execution time, IPC and speedup for different optimizations of \texttt{susan\_c}, averaged across a subset of the
    inputs. We use correlation coefficient~\cite{Bishop:2006:PRM:1162264} to quantify if  IPC and execution time are strongly correlated to
    speedup.
    %This correlation coefficient takes a value between -1 and 1, the closer the coefficient is to +/-1, the stronger the
    %correlation between the variables.
    In our case, the IPC and execution time show correlation coefficients of $-0.14$ and $-0.16$,
    respectively, to speedup.
    %These values are far from 1, suggesting that there is a weak correlation between speedup and IPC or execution time.
    Neither IPC nor execution time correlate well, and so neither can be used for iterative compilation. We need a work efficiency
    metric which can approximate speedup instead, using only a single execution.

### 2017-09-15 prof.tex
overall changes

\section{Work Profiling}

In this section, we describe how to measure work online. Total work is just the sum of the costs of all executed basic blocks, so it is
possible to embed its computation into the execution of the program. A na\"{\i}ve instrumentation would use a global counter initialized at
the interception value, $\varepsilon$. Each basic block would then have only to increment this counter with the total cost of its
instructions every time it is executed. Although easy to implement, the na\"{\i}ve approach introduces a significant overhead.

@@

\paragraph{Populating the edge flows:}
Intuitively, if all the edge flows are known for the complement of a spanning tree then at any leaf of the spanning tree there is only one
unknown edge flow. This unknown edge flow can be calculated by Kirchhoff's first law\cite{knuth73,ball94}. This process repeats as a bottom-up
propagation until all the unknown edge flows have been calculated. This algorithm can be formally defined as a post-order traversal on the
spanning tree. Let $G$ be the CFG. This algorithm first initializes the edge flows $D_{(u,v)}$, for each edge $(u,v)$ in the CFG, as
follows:
\[
D_{(u,v)} \gets
\begin{cases}
    P_{(u,v)} & \quad \text{if edge $(u,v)$ has a probe $P_{(u,v)}$}\\
    0       & \quad \text{otherwise}
\end{cases}
\]
Afterwards, for each vertex $u\in G$, in a post-order traversal of the spanning tree,
we can use the Kirchhoff's first law and apply the following operations
in a symbolic fashion:
\[
\Sigma^+_u = \sum_{v\in N^+(u)} D_{(u,v)}
\]
\[
\Sigma^-_u = \sum_{v\in N^-(u)} D_{(v,u)}
\]
\[
\forall v\in N^+(u):  D_{(u,v)} \gets
\begin{cases}
    D_{(u,v)} & \quad \text{if $D_{(u,v)}\neq 0$}\\
    \Sigma^-_u - \Sigma^+_u       & \quad \text{otherwise}
\end{cases}
\]
\[
\forall v\in N^-(u):  D_{(v,u)} \gets
\begin{cases}
    D_{(v,u)} & \quad \text{if $D_{(v,u)}\neq 0$}\\
    \Sigma^+_u - \Sigma^-_u       & \quad \text{otherwise}
\end{cases}
\]

>>>>

\section{Work Profiling}\label{sec:prof}

In this section, we describe how to measure work online. Total work is just the sum of the costs of all executed basic blocks, so it is
possible to embed its computation into the execution of the program. A na\"{\i}ve instrumentation would use a global counter initialized 
to zero. Each basic block would then have only to increment this counter with the total cost of its
instructions every time it is executed. Although easy to implement, the na\"{\i}ve approach introduces a significant overhead.

@@

\paragraph{Populating the edge flows:}
Intuitively, if all the edge flows are known for the complement of a spanning tree then at any leaf of the spanning tree there is only one
unknown edge flow. This unknown edge flow can be calculated by Kirchhoff's first law\cite{knuth73,ball94},
also known as the principle of \textit{conservation of flow}, which states that
the amount of flow into a vertex equals the amount of output flow.
This process repeats as a bottom-up propagation until all the unknown edge flows have been calculated. This algorithm can be formally defined as a post-order traversal on the
spanning tree. Let $G$ be the CFG. This algorithm first initializes the edge flows $D_{(u,v)}$, for each edge $(u,v)$ in the CFG, as
follows:
\[
D_{(u,v)} \gets
\begin{cases}
    P_{(u,v)} & \quad \text{if edge $(u,v)$ has a probe $P_{(u,v)}$}\\
    0       & \quad \text{otherwise}
\end{cases}
\]
This initialization represents only the trivially known edge flows, which are the
probes themselves (the highlighted edges in red).

Afterwards, the algorithm consists in propagating the known edge flows, infering
the symbolic expressions in the unknown edge flows.
This is achieved by performing a bottom-up propagation from the trivially known
edge flows (in red) by applying the Kirchhoff's first law, i.e.,
for each vertex with a single unknown edge flow, we compute the difference between
the known incoming edge flows by the known outgoing edge flows.
Formally,
for each vertex $u\in G$, in a post-order traversal of the spanning tree,
we can use the Kirchhoff's first law and apply the following operations
in a symbolic fashion:
\[
\Sigma^+_u = \sum_{v\in N^+(u)} D_{(u,v)}
\]
\[
\Sigma^-_u = \sum_{v\in N^-(u)} D_{(v,u)}
\]
\[
\forall v\in N^+(u):  D_{(u,v)} \gets
\begin{cases}
    D_{(u,v)} & \quad \text{if $D_{(u,v)}\neq 0$}\\
    \Sigma^-_u - \Sigma^+_u       & \quad \text{otherwise}
\end{cases}
\]
\[
\forall v\in N^-(u):  D_{(v,u)} \gets
\begin{cases}
    D_{(v,u)} & \quad \text{if $D_{(v,u)}\neq 0$}\\
    \Sigma^+_u - \Sigma^-_u       & \quad \text{otherwise}
\end{cases}
\]

### 2017-09-15 work.tex
overall changes

    Similarly to previous work~\citep{giusto01,powell09,brandolese11}, the coefficients, $w_i$, are found by linear regression. We create
    one regressor data point for each of a set of program and input pairs. The programs are given in Table~\ref{tab:kdatasets:training},
    and each has 1,000 inputs. The response variable is the runtime of the unoptimized program as it processes the input. To ensure
    statistically sound results, the runtime is measured by repeated execution until the 99\% confidence interval is no larger than 1\% of
    the mean. The explanatory variables are the dynamic instruction counts for each program and input pair. They are recorded by a separate,
    instrumented run of the program on the input.

    Figure~\ref{fig:cost-model} shows the correlation of the model with runtime on instances of test benchmarks. The correlation
    coefficient is \FIXME{XXX}, and the mean absolute percentage error (MAPE) is \FIXME{XXX}. Figure~\ref{fig:motivation-work-efficiency}
    shows a similar plot to Figure~\FIXME{\ref{fig:???}}, of work efficiency versus speedup for 500 different binaries of \texttt{susan\_c}. Here the new metric correlates much better than either performance or IPC.

    \begin{figure}[t]
        \centering
        \includegraphics[width=0.6\linewidth]{figs/cost-model.pdf}
        \caption{
            Linear model for work metric fitted from empirical data.
            \red{zw: bigger font size.}
        }
        \label{fig:cost-model}
    \end{figure}

    \begin{figure}[t]
        \centering
        \includegraphics[width=0.5\textwidth]{figs/motivation-work-efficiency.pdf}
        \caption{Relationship between work efficiency vs speedup for 500 different binaries of \texttt{susan\_c}.
                Each point represents averages over a subset of all inputs.}
        \label{fig:motivation-work-efficiency}
    \end{figure}

>>>>

    Similarly to previous work~\citep{giusto01,powell09,brandolese11}, the coefficients, $w_i$, are estimated by linear regression. We create
    one regressor data point for each of a set of program and input pairs.
    We use a separate set of programs from those discussed in Sections~\ref{sec:setup} and \ref{sec:results}.
    %for estimating the architectural settings for the work metric.
    The response variable is the execution time of the unoptimized program as it processes the input. To ensure
    statistically sound results, the execution time is measured by repeated execution until the 99\% confidence interval is no larger than 1\% of
    the mean. The explanatory variables are the dynamic instruction counts for each program and input pair. They are recorded by a separate,
    instrumented run of the program on the input.

    Figure~\ref{fig:work-vs-O0time} shows the correlation of the work metric ($\Delta W$) with runtime on instances of the unoptimized \texttt{susan\_c} benchmark.
    Each point represents the execution of a single input, from a total of 1,000 inputs.
    The correlation coefficient is $0.99$, and the mean absolute percentage error (MAPE) is only about 3.5\%.
    This result suggests that the new metric correlates much better than either execution time or IPC.

    %Figure~\ref{fig:motivation-work-efficiency} shows a similar plot to Figure~\FIXME{\ref{fig:motivation-metric}},
    %of work efficiency versus speedup for 500 different binaries of \texttt{susan\_c}.
    %Here the new metric correlates much better than either execution time or IPC,
    %exhibiting a correlation coefficient of $0.93$ with speedup.

    \begin{figure}[t]
        \centering
        \includegraphics[width=0.5\textwidth]{figs/work-correlation-with-O0_susan_c.pdf}
        \caption{
            Relationship between work metric ($\Delta W$) and the execution time
            of the unoptimized version of the \texttt{susan\_c} benchmark over the
            1,000 inputs.
            Each point represents the execution of a single input.
            They have a correlation coefficient of $0.99$.
        }
        \label{fig:work-vs-O0time}
    \end{figure}

    %\begin{figure}[t]
    %    \centering
    %    \includegraphics[width=0.6\linewidth]{figs/cost-model.pdf}
    %    \caption{
    %        Linear model for work metric fitted from empirical data.
    %        \red{zw: bigger font size.}
    %    }
    %    \label{fig:cost-model}
    %\end{figure}

    %\begin{figure}[t]
    %    \centering
    %    \includegraphics[width=0.5\textwidth]{figs/motivation-work-efficiency.pdf}
    %    \caption{Relationship between work efficiency \textit{vs} speedup for 500 different binaries of \texttt{susan\_c}.
    %            Each point represents the average over a subset of all inputs.
    %            The correlation coefficient is $0.93$.}
    %    \label{fig:motivation-work-efficiency}
    %\end{figure}
